---
title: "HW3"
author: "Peiran Chen"
date: "4/30/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
library(ISLR2)
library(knitr)
library(boot)
```

## 1.

### a.

```{r ex1a}
set.seed(123)
y <- rnorm(100)
x <- matrix(rnorm(10000*100), ncol=10000)

lm_1a <- lm(y ~ x)
```

We can write our expression as:


\begin{align*}
Y_i &= \beta_{0,i} + \beta_{1,i} X_{1,i} + \beta_{2,i}X_{2,i} + .... + \beta_{10000,i}X_{10000,i} + \varepsilon_i, \text{ where } \varepsilon \sim N(0,1) \\
\intertext{In matrix algebra form, we have:}\\
Y &= \beta X + \varepsilon \\  
\intertext{And Least Squares will give us}\\  
\hat Y &= \hat\beta X\\
\intertext{Hence, it's MSE is}\\
MSE &= E[Y - \hat Y]^2\\
&= Var[\varepsilon] + Bias^2[\hat f(x_0)] + Var[\hat f(x_0)]
\end{align*}


### b.

As we can see, that $Var[\varepsilon]$ is irreducible, because it exists in nature,
and we know that it is caused by the randomness in our $y$, which is just $1$. 


### c.




#### i.

$$
Bias = \sqrt{\left(E[\hat f(x_0)] - f(x_0)\right)^2} = f(x_0) 
$$
Because we have $f(x) = y \sim N(0,1)$. Thus, we have the Bias is equal to 0.


#### ii.

```{r ex1cii}
fc <- rep(0,100)
variance <- (sd(fc))^2
```

$$
Var[\hat f(x)] = Var[0] = 0
$$
Thus, the variance for this procedure is `r variance`.

#### iii.

In this case, 

$$
EPE =  Var[\varepsilon] + Bias^2[\hat f(x_0)] + Var[\hat f(x_0)] = Var[\varepsilon] = 1
$$
And our $Bias^2[\hat f(x)] = 0$, and $Var[\hat f(x)] = 0$. And $Var[\varepsilon] = 1$, we have \textbf{EPE} = 1.

#### iv.

With Validation Set Approach, we first scramble the data set into train and validation, and
then, then, with $\hat f_{\text{train}}(\text{validation} \ x) = 0$. It's 
\[CV_{VSA} = \frac1n\sum_{i \in \text{validation set}}(\hat f_{\text{train}}(x_i) - y_i)^2 = E[y_i^2] = Var[y_i] + E[y_i]^2 = 1\]. 

```{r ex1iv, warning = FALSE}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]
lm_2 <- lm(train.y ~ train.x)
mean((0 - predict.lm(lm_2, data.frame(val.x)))^2)
```
And it is close to our derivation for $CV_{VSA}$.

#### v.

I found out that they do match with each other. And the reason for that is because we have
set that $$\hat f(x) = 0 \ \forall x \in \mathbb{R} $$. Hence, it will has Variance of 0 and Bias
of 0.


### d.

```{r ex1d, warning = FALSE, message=FALSE}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]
lm_2 <- lm(train.y ~ train.x)
mean((val.y - predict.lm(lm_2, data.frame(val.x)))^2)
```


### e.

(c) has a smaller Estimated Test Error, along with 0 Bias and 0 Variance, which are
definitely lower than part (d)'s Bias and Variance. It might seems like that (c) is
a better choice. And it all comes down to how we generated our data in the first place.
We generated $y$ and $x$ independently with the same $Normal(0,1)$ distribution,
so in a perfect setting, their covariance would be


\begin{align*}
Cov[X,Y] &= E[XY]-E[X]E[Y]\\
&= E[X]E[Y] - E[X]E[Y]\\
&= 0\\
\\
Cor[X, Y] &= \frac{Cov[X, Y]}{\sqrt{Var[X]Var[Y]}}\\
&= 0
\end{align*}


And we are doing least squares to find relationship on this 0 correlation relationship in (d).
Hence, (c)'s choice of $\hat f(x) = 0 \forall x$ seems to be a reasonable and good choice.
And that answers why (c) gives us a lower estimation test error when comparing to (d). And my
asnwer does support this.


## 2.

### a.

```{r ex2a}
hist(cor(x[,c(1:10000)],y))

df <- tibble("index" = c(1:10000),
             "Correlation" = abs(cor(x[,c(1:10000)],y)))

top_10 <- top_n(df, 10, Correlation)
```

### b.

```{r ex2b}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]

lm_3 <- lm(train.y ~ train.x[,top_10$index])
mean((val.y - predict.lm(lm_3, data.frame(val.x[,top_10$index]))) ^ 2)
```

### c.
```{r ex2c}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]

df_new <- tibble("index" = c(1:10000),
             "Correlation" = cor(train.x[,c(1:10000)],train.y))

top_10_new <- top_n(df_new, 10, Correlation)

lm_4 <- lm(train.y ~ train.x[,top_10_new$index])
mean((val.y - predict.lm(lm_4, data.frame(val.x[,top_10_new$index]))) ^ 2)
```

### d.

In our finding, we found out that with the same $q = 10$ features, "Option 1" gave
us a lower estimated test error when compare to "Option 2". And the reason for that
is when selecting top 10 predictors, we used all the information/data we have for
"Option 1", and that includes validation set, which is not appropriate when performing
subset selection. Hence, we shall avoid using this and this way of modeling does not 
give us meaningful answers. Hence, despite having higher test error, we shall use 
"Option 2" to perform this task.

## 3.

### a.

I choose a data set that is called "Superconductivty Data Data Set". It has $p = 81$ features. And the goal is to use all the features(extracted from the superconductor’s chemical formula) to predict the superconducting critical temperature.


### b.

```{r ex3b}
df <- read.csv("train.csv")

set.seed(123)
# We would first split the data into train & val sets
train <- sample(21263, 10632)

df_train <- df[train, ]
df_val <- df[-train, ]

lm_5 <- lm(critical_temp ~.-critical_temp, data = df_train)
test_error <- mean((df_val$critical_temp - predict.lm(lm_5, df_val[1:81])) ^ 2)
```

The test error(Test MSE) is `r test_error`. And I first use Validation Set Approach
to split the data into train/validation set about half and half(10632:10631). Then,
I perform Least Squares on the training data. And calculate the Test Error using
validation set data.

### c.

```{r ex3c}
lambdas <- 10^seq(10, -2, length=100)
ridge_mod <- glmnet(as.matrix(df_train[1:81]), as.matrix(df_train[82]), alpha = 0, lambda = lambdas)

dim(coef(ridge_mod))

plot(ridge_mod, xvar="lambda")
```

### d.

```{r ex3d}
cv.out <- cv.glmnet(as.matrix(df[1:81]), as.matrix(df[82]), alpha = 0)
plot(cv.out)

best_lambda <- cv.out$lambda.min

smallest_test_error <- min(cv.out$cvm)
```

When $\lambda =$ `r best_lambda`, we will have the smallest test error of `r smallest_test_error`.


### e.

```{r ex3e}
lambdas <- 10^seq(10, -2, length=100)
lasso_mod <- glmnet(as.matrix(df_train[1:81]), as.matrix(df_train[82]), alpha = 1, lambda = lambdas)

dim(coef(lasso_mod))

plot(lasso_mod, xvar="lambda")
```

### f.

```{r ex3f}
cv.out <- cv.glmnet(as.matrix(df[1:81]), as.matrix(df[82]), alpha = 1)
plot(cv.out)

best_lambda <- cv.out$lambda.min

smallest_test_error <- min(cv.out$cvm)

coef(cv.out, s=best_lambda)
```

Best lambda in this case is `r best_lambda`, we will have the smallest test error of `r smallest_test_error`. As we can see from above, this time, LASSO model uses all features except "entropy_atomic_radius", "mean_Valence", and "std_Valence".


## 4.

### a.

```{r ex4a}
data(Auto)
# Split data
set.seed(434)
degrees <- c(1:10)
train <- sample(392, 196)
test_MSE <- matrix(NA, 10, 10)
for(i in degrees){
  lm.fit <- lm(mpg ~ poly(horsepower, i), data = Auto, subset = train)
  test_MSE[1, i] <- mean((Auto$mpg - predict(lm.fit, Auto))[-train] ^ 2)
}

plot(x = degrees, y = test_MSE[1, ],
     lwd = 2,
     type = "l",
     lty = 1,
     ylim = c(15,27),
     xlab = "Degree of Polynomial",
     ylab = "Mean Squared Error")


for (i in 2:10) {
  train <- sample(392, 196)
  for (j in degrees) {
    lm.fit <- lm(mpg ~ poly(horsepower, j),
                 data = Auto,
                 subset = train)
    test_MSE[i, j] <- mean((Auto$mpg - predict(lm.fit, Auto))[-train] ^ 2)
  }
  lines(x = degrees, y = test_MSE[i, ], col = i,
        lty = 1,
        lwd = 2)
}

kable(tibble(
  Degree = degrees,
  "Mean Validation MSE" = rowMeans(test_MSE)
))

```

As we can see from the above table, Mean Test MSE is the smallest when degree is at 7, with 
a value of 16.25580.  
I found out that the test error is \textbf{not} monotonically decreasing. And after 
a sharp decrease in Degree 2, we see little to no improvements, and sometimes even
worse test error with higher degrees.

### b.
```{r ex4b}
# LOOCV
set.seed(435)
cv_error <- rep(NA, 10)

for( i in 1:10){
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv_error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}

kable(tibble(
  Degree = degrees,
  "LOOCV" = cv_error
))

plot(x = degrees,
     y = cv_error,
     xlab = "Degree of Polynomial",
     ylab = "Mean Squared Error",
     main = "LOOCV",
     col = "purple",
     type = "b",
     pch = 16)
```


With LOOCV, the lowest error occurs at Degree equals 7, with mean test error of 18.83305.  
I found out that the test error is \textbf{not} monotonically decreasing. And after 
a sharp decrease in Degree 2, we see little to no improvements, and sometimes even
worse test error with higher degrees.

### c.

```{r ex4c}
set.seed(435)

cv_error_10 <- matrix(0, 10, 10)

for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv_error_10[1, i] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
plot(
  x = c(1:10),
  y = cv_error_10[1,],
  lty = 1,
  type = "l",
  lwd = 2,
  ylim = c(16, 28),
  xlab = "Degree of Polynomial",
  ylab = "Mean Squared Error",
  main = "10-fold CV"
)


for (i in 2:10) {
  for (j in 1:10) {
    glm.fit <- glm(mpg ~ poly(horsepower, j), data = Auto)
    cv_error_10[i, j] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
  }
  lines(x = c(1:10), y = cv_error_10[i, ],
        col = i,
        lwd = 2)
}

kable(tibble(
  Degree = degrees,
  "Mean Validation MSE" = rowMeans(cv_error_10)
))
```

At degree equals to 5, we have the smallest mean 10 fold CV error of 19.53472. 
I found out that the test error is \textbf{not} monotonically decreasing. And after 
a sharp decrease in Degree 2, we see little to no improvements, and sometimes even
worse test error with higher degrees.


### d.

```{r ex4d}
set.seed(435)
degrees <- c(1:10)
train_MSE <- rep(NA, 10)

for(i in degrees){
  lm.fit <- lm(mpg ~ poly(horsepower, i), data = Auto)
  train_MSE[i] <- mean(lm.fit$residuals^2)
}

plot(x = degrees,
     y = train_MSE,
     xlab = "Degree of Polynomial",
     ylab = "Training Set Mean Squared Error",
     main = "Linear Model with 1-10 degree of Polynomial",
     col = "red",
     type = "b",
     pch = 16)

kable(tibble(
  Degree = degrees,
  "Mean Validation MSE" = train_MSE
))
```

With increase in Degree of Polynomial, our Training MSE will monotonically decrease, and
the smallest Train MSE occurred at $degree = 10$. 

### e.

```{r ex4e}
set.seed(435)

lm.fit <- lm(mpg ~ poly(horsepower, 10), data = Auto)
summary(lm.fit)
```

As we can see from summary table, the intercept, first, and second degree
polynomials are statistically significant for having p-value<0.001. And Degree 5,
p-value<0.01, Degree 6, p-value < 0.05, Degree 7, p-value < 0.1. However, after Degree
2, all the rest are minor changes/improvement to our test error. Hence I would say
using 2 degree of freedom polynomial fit is a good enough practice. And it does match the
graph we plotted from (a)-(d), where we saw a sharp decrease in Test/Train MSE at 2 degree
of polynomial.


## 5.


### a.



\begin{align*}
\arg\min_\beta \sum_{i = 1}^n (y_i - \beta x_i)^2 = \frac{\partial}{\partial\beta}\sum_{i = 1}^n (y_i - \beta x_i)^2 &= 0\\
-2 \sum_{i = 1}^n x_i(y_i - \beta x_i) &= 0\\
\sum_{i = 1}^n x_iy_i - \beta\sum_{i = 1}^n x_i^2&= 0\\
\hat\beta^{L.S.} = \frac{\sum_{i = 1}^n x_iy_i}{\sum_{i = 1}^n x_i^2}
\end{align*}


### b.


\begin{align*}
\arg\min_\beta \sum_{i = 1}^n (y_i - \beta x_i)^2 + \lambda\beta^2 = \frac{\partial}{\partial\beta}\left[\sum_{i = 1}^n (y_i - \beta x_i)^2 + \lambda\beta^2\right] &= 0\\
-2 \sum_{i = 1}^n x_i(y_i - \beta x_i)  + 2\lambda\beta&= 0\\
\sum_{i = 1}^n x_iy_i - \beta\sum_{i = 1}^n x_i^2 &= \lambda\beta\\
\beta\left(\sum_{i = 1}^n x_i^2 + \lambda\right) &= \sum_{i = 1}^n x_iy_i\\
\hat\beta^{L.S.} &= \frac{\sum_{i = 1}^n x_iy_i}{\sum_{i = 1}^n x_i^2 + \lambda}\\
\end{align*}



### c.

From part a),


\begin{align*}
E[\hat\beta^{L.S.}] &= E\left[\frac{\sum_{i = 1}^n x_iy_i}{\sum_{i = 1}^n x_i^2}\right] \\
&= E\left[\frac{\sum_{i = 1}^n x_i(3x_i + \varepsilon))}{\sum_{i = 1}^n x_i^2}\right] \\
&= E\left[\frac{\sum_{i = 1}^n 3x_i^2 + \varepsilon \sum_{i = 1}^nx_i}{\sum_{i = 1}^n x_i^2}\right] \\
\intertext{Since } X \text{ and } \varepsilon \text{ are independent}\\
&= \frac{3E[\sum_{i = 1}^n x_i^2] + E[\varepsilon] \times E[\sum_{i = 1}^nx_i]}{E[\sum_{i = 1}^n x_i^2]} \\
&= \frac{3E[\sum_{i = 1}^n x_i^2] }{E[\sum_{i = 1}^n x_i^2]} \\
&= 3 = \beta
\end{align*}


Hence, $\hat\beta$ is an unbiased estimator of $\beta$.

### d.


\begin{align*}
E[\hat\beta^{ridge}] &= E\left[\frac{\sum_{i = 1}^n x_iy_i}{\sum_{i = 1}^n x_i^2 + \lambda}\right]\\
&= E\left[\frac{\sum_{i = 1}^n x_i(3x_i + \varepsilon))}{\sum_{i = 1}^n x_i^2+ \lambda}\right] \\
&= E\left[\frac{\sum_{i = 1}^n 3x_i^2 + \varepsilon \sum_{i = 1}^nx_i}{\sum_{i = 1}^n x_i^2+ \lambda}\right] \\
\intertext{Since } X \text{ and } \varepsilon \text{ are independent}\\
&= \frac{3E[\sum_{i = 1}^n x_i^2] + E[\varepsilon] \times E[\sum_{i = 1}^nx_i]}{E[\sum_{i = 1}^n x_i^2] + E[\lambda]} \\
&= \frac{3E[\sum_{i = 1}^n x_i^2]}{E[\sum_{i = 1}^n x_i^2] + E[\lambda]} \\
\intertext{Since x is fixed, and lambda is a constant}\\
&= \frac{3\sum_{i = 1}^n x_i^2}{\sum_{i = 1}^n x_i^2 + \lambda} \\
\end{align*}

Hence, we have shown that Ridge regression will give us a biased estimator $\hat\beta^{ridge}$ for $\beta$. However, as $\lambda \to 0$, we will have an less biased estimator for $\beta$.
And finally when $E[\lambda] = 0$, $\hat\beta^{ridge}$ will be unbiased.


### e.


\begin{align*}
Var[\hat\beta^{L.S.}] &= E\left[\hat\beta - E[\hat\beta]\right]^2\\
&=E[\hat\beta - \beta]^2\\
&=E\left[\frac{\sum_{i = 1}^n 3x_i^2 + \varepsilon \sum_{i = 1}^nx_i}{\sum_{i = 1}^n x_i^2} - 3\right]^2 \\
&=E\left[\frac{\varepsilon \sum_{i = 1}^nx_i}{\sum_{i = 1}^n x_i^2} \right]^2 \\
&= \frac{1}{\sum_{i = 1}^n x_i^4}E\left[ \left(\sum_{i = 1}^n\varepsilon x_i\right)^2\right]\\
&= \frac{1}{\sum_{i = 1}^n x_i^4}E\left[ \left(x_1\varepsilon + x_2\varepsilon + ... + x_n\varepsilon\right)^2\right]\\
&= \frac{1}{\sum_{i = 1}^n x_i^4}E\left[x_1^2\varepsilon^2 + x_2^2\varepsilon^2 + ... + x_n^2\varepsilon^2 + 2x_1x_2\varepsilon^2+... + 2x_{n-1}x_n\varepsilon^2\right]\\
&= \frac{E[\varepsilon^2]}{\sum_{i = 1}^n x_i^4} E[x_1^2 + x_2^2+ ... + x_n^2 + 2x_1x_2+... + 2x_{n-1}x_n]\\
\text{Because we are given } &Cov(\varepsilon_i, \varepsilon_i') =  0 \ \forall i \neq i'\\
&= \frac{E[\varepsilon^2]}{\sum_{i = 1}^n x_i^4} E[x_1^2 + x_2^2+ ... + x_n^2]\\
&= \frac{Var[\varepsilon] + E[\varepsilon]^2}{\sum_{i = 1}^n x_i^4}\sum_{i = 1}^n x_i^2\\
&= \frac{\sigma^2}{\sum_{i = 1}^n x_i^2}
\end{align*}




### f.


\begin{align*}
Var[\hat\beta^{ridge}] &= E\left[\hat\beta^{ridge} - E[\hat\beta^{ridge}]\right]^2\\
&=E\left[\frac{\sum_{i = 1}^n 3x_i^2 +  \sum_{i = 1}^n \varepsilon_ix_i}{\sum_{i = 1}^n x_i^2 + \lambda} - \frac{3\sum_{i = 1}^n x_i^2}{\sum_{i = 1}^n x_i^2 + \lambda}\right]^2\\
&=E\left[\frac{ \sum_{i = 1}^n \varepsilon_ix_i}{\sum_{i = 1}^n x_i^2 + \lambda} \right]^2 \\
&= \frac{1}{(\sum_{i = 1}^n x_i^2 + \lambda)^2}E\left[ \left(\sum_{i = 1}^n\varepsilon x_i\right)^2\right]\\
&= \frac{1}{(\sum_{i = 1}^n x_i^2 + \lambda)^2}E\left[ \left(x_1\varepsilon + x_2\varepsilon + ... + x_n\varepsilon\right)^2\right]\\
&= \frac{1}{(\sum_{i = 1}^n x_i^2 + \lambda)^2}E\left[x_1^2\varepsilon^2 + x_2^2\varepsilon^2 + ... + x_n^2\varepsilon^2 + 2x_1x_2\varepsilon^2+... + 2x_{n-1}x_n\varepsilon^2\right]\\
&= \frac{E[\varepsilon^2]}{(\sum_{i = 1}^n x_i^2 + \lambda)^2} E[x_1^2 + x_2^2+ ... + x_n^2 + 2x_1x_2+... + 2x_{n-1}x_n]\\
\text{Because we are given } &Cov(\varepsilon_i, \varepsilon_i') =  0 \ \forall i \neq i'\\
&= \frac{E[\varepsilon^2]}{(\sum_{i = 1}^n x_i^2 + \lambda)^2} E[x_1^2 + x_2^2+ ... + x_n^2]\\
&= \frac{Var[\varepsilon] + E[\varepsilon]^2}{(\sum_{i = 1}^n x_i^2 + \lambda)^2}\sum_{i = 1}^n x_i^2\\
&= \frac{\sigma^2}{(\sum_{i = 1}^n x_i^2 + \lambda)^2}\sum_{i = 1}^n x_i^2
\end{align*}



### g.

So we have


\begin{align*}
Bias[\hat\beta^{ridge}] &= E[\hat\beta^{ridge}] - \beta\\
&= \frac{3\sum_{i = 1}^n x_i^2}{\sum_{i = 1}^n x_i^2 + \lambda} - 3\\
&= 3\left(\frac{\sum_{i = 1}^n x_i^2}{\sum_{i = 1}^n x_i^2 + \lambda} - \frac{\sum_{i = 1}^n x_i^2 + \lambda}{\sum_{i = 1}^n x_i^2 + \lambda}\right)\\
&= 3 \frac{-\lambda}{\sum_{i = 1}^n x_i^2 + \lambda}\\
Bias^2[\hat\beta^{ridge}] &= 9\frac{\lambda^2}{(\sum_{i = 1}^n x_i^2 + \lambda)^2}\\
\end{align*}


And suppose we have $\sum_{i = 1}^n x_i^2  = 10$, $\sigma = 1$.

```{r ex4g}
x <- 10
lambda <- seq(0, 5, by = 0.001)
sigma <- 1
plot(lambda, (-3*lambda/(x + lambda))^2, type = "l", col = "purple",
     ylab = "Value", ylim = c(-1,1))
lines(lambda, -3*lambda/(x + lambda), type = "l", col = "red")
lines(lambda, 10*(sigma^2/(10 + lambda)^2), col = "blue")
legend("topleft",
       c("Squared Bias", "Bias", "Variance"),
       col = c("purple", "red", "blue"),
       lty = 1)
```

As we can see from above, as $lambda$ increases, $Bias^2$ increases and takes the value 0 at 0. And it's not hard to see that $\hat\beta^{ridge} = \hat\beta^{L.S.}$ if $\lambda = 0$.(This is the reason why Bias/Squared Bias is 0 at 0). Then, as $\lambda$ goes to $9$, since $\lambda \to \infty$ shows that the ridge estimate would be 0 and hence bias would be $3$. For variance though, we see that increase $\lambda$ reduce variances. Because larger penalty that introduced by $\lambda$ will force the weight to shrink towards zero and reduce the scale and variance. Thus, we will see that larger penalty in ridge-regression increases the Squared Bias for the estimate and reduces the variance. a.k.a, Bias-Variance Trade-off.


