---
title: "HW3"
author: "Peiran Chen"
date: "4/30/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
```

## 1.

### a.

```{r ex1a}
set.seed(123)
y <- rnorm(100)
x <- matrix(rnorm(10000*100), ncol=10000)

lm_1a <- lm(y ~ x)
```

In matrix form, we could rewrite our expression as:

$$
\begin{align*}
Y &= \beta_0 + \beta_1 X_1 + \beta_2X_2 + .... + \beta_{10000}X_{10000} + \varepsilon, \text{ where } \varepsilon \sim N(0,1) \\
Y &= \beta X + \varepsilon \\  
\intertext{And Least Squares will give us}\\  
\hat Y &= \hat\beta X + \varepsilon\\
\intertext{Hence, it's MSE is}\\
MSE &= E[Y - \hat Y]^2\\
&= Var[\varepsilon] + Bias^2[\hat f(x_0)] + Var[\hat f(x_0)]
\end{align*}
$$

### b.

As we can see, that $Var[\varepsilon]$ is irreducible, because it exists in nature,
and we know that it is caused by the randomness in our $y$, which is just $1$. 


### c.




#### i.

$$
Bias = \sqrt{\left(E[\hat f(x)] - f(x)\right)^2} = \sqrt{f(x)^2} =f(x)
$$
Because we have $f(x) = y \sim N(0,1)$. Thus, we have the Bias is equal to 0.


#### ii.

```{r}
fc <- rep(0,100)
(variance <- (sd(fc))^2)
```

Thus, the variance for this procedure is `r variance`.

#### iii.

In this case, 

$$
EPE =  Var[\varepsilon] + Bias^2[\hat f(x_0)] + Var[\hat f(x_0)] = Var[\varepsilon] = 1
$$
And our $Bias^2[\hat f(x)] = 0$, and $Var[\hat f(x)] = 0$. And $Var[\varepsilon] = 1$, we have \textbf{EPE} = 1.

#### iv.

With Validation Set Approach, we first scramble the data set into train and validation, and
then, then, with $\hat f_{\text{train}}(\text{validation} \ x) = 0$. It's 
\[CV_{VSA} = \frac1n\sum_{i \in \text{validation set}}(\hat f_{\text{train}}(x_i) - y_i)^2 = E[y_i^2] = Var[y_i] + E[y_i]^2 = 1\]. 


#### v.

I found out that they do match with each other. And the reason for that is because we have
set that $\hat f(x) = 0 \ \forall x \in \mathbb{R} $. 

### d.

```{r, warning = FALSE}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]
lm_2 <- lm(train.y ~ train.x)
mean((val.y - predict.lm(lm_2, data.frame(val.x))) ^ 2)
```


### e.

(d) has a smaller estimated Test Error, but higher variance and bias.


## 2.

### a.

```{r}
hist(cor(x[,c(1:10000)],y))

df <- tibble("index" = c(1:10000),
             "Correlation" = abs(cor(x[,c(1:10000)],y)))

top_10 <- top_n(df, 10, Correlation)
```

### b.

```{r}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]

lm_3 <- lm(train.y ~ train.x[,top_10$index])
mean((val.y - predict.lm(lm_3, data.frame(val.x[,top_10$index]))) ^ 2)
```

### c.
```{r}
set.seed(123)
train <- sample(100, 50)
train.x <- x[train, ]
val.x <- x[-train, ]

train.y <- y[train]
val.y <- y[-train]

df_new <- tibble("index" = c(1:10000),
             "Correlation" = cor(train.x[,c(1:10000)],train.y))

top_10_new <- top_n(df_new, 10, Correlation)

lm_4 <- lm(train.y ~ train.x[,top_10_new$index])
mean((val.y - predict.lm(lm_4, data.frame(val.x[,top_10_new$index]))) ^ 2)
```

## 3.

### a.

I choose a data set that is called "Superconductivty Data Data Set". It has $p = 81$ features. And the goal is to use all the features(extracted from the superconductor’s chemical formula) to predict the superconducting critical temperature.


### b.

```{r}
df <- read.csv("train.csv")

set.seed(123)
# We would first split the data into train & val sets
train <- sample(21263, 10632)

df_train <- df[train, ]
df_val <- df[-train, ]

lm_5 <- lm(critical_temp ~.-critical_temp, data = df_train)
test_error <- mean((df_val$critical_temp - predict.lm(lm_5, df_val[1:81])) ^ 2)
```

The test error(Test MSE) is `r test_error`. And I first use Validation Set Approach
to split the data into train/validation set about half and half(10632:10631). Then,
I perform Least Squares on the training data. And calculate the Test Error using
validation set data.

### c.

```{r ex3c}
lambdas <- 10^seq(10, -2, length=100)
ridge_mod <- glmnet(as.matrix(df_train[1:81]), as.matrix(df_train[82]), alpha = 0, lambda = lambdas)

dim(coef(ridge_mod))

plot(ridge_mod, xvar="lambda")
```

### d.

```{r}
smallest <- 1000
target <- 0
for (i in lambdas){
  ridge_pred <- predict(ridge_mod, s = i, newx=as.matrix(df_val[1:81]))
  error <- mean((ridge_pred - as.matrix(df_val[82]))^2)
  if (error < smallest){
    smallest <- error
    target <- i
  }
}



```

When $\lambda = $

## 4.

### a.

```{r}

```

