---
title: "STAT 435 HW1"
author: "Peiran Chen"
date: "3/31/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.

### a) 

Taking a parametric approach will have following pros:  
 - Does not need a lot of data  
 - Simplifies the problem because it is generally much easier to estimate a set of parameters
and cons:  
 - The model we choose will usually not match the true unknown form of $f$, and if
 the chosen model is too far from the true $f$, then our estimate will be poor.  
Taking a nonparametric approach will have following pros:  
 - Avoid unnecessary assumptions about the functional form of $f$ will have the
 potential to accurately fit a wider range of possible shapes for $f$.  
 and cons:  
 - A very large number of observations is required in order to obtain an accurate
 estimate for $f$.  
 
### b)

For parametric approach, I would say when we have a small number of observations
to work with, such as getting survey on people's blood pressure and hours of physical 
exercises they do each week. And we know that having more time to exercises will result in a lower blood pressure
as a matter of fact. Hence we can make assumptions to $f$ in this case to be a 
linear model:
\[blood\;pressure \approx \beta_0 + \beta_1 \times physical\;exercises\]

### c)

For non-parametric approach, I would say when we have a lot of data to work with,
we can use this method to do the same prediction as part b). 



## 2.

### a) 

In this case, I would expect the inflexible methods to perform better. Since sample
size is small, there won't be enough data for flexible methods such as deep learning
and etc. Also, the number of predictors are large, hence it would be a good practice
to use OLS so that we have more interpretability. Hence inflexible methods tends to perform better.

### b)

In this case, I would expect the flexible method to perform better.


## 3.

### a)
 It is a regression problem. And the goal is prediction, where $n = 50, p = 8$.
 
### b)
 It is a classification problem. And the goal is inference, where $n = 50, p = 6$.
 
## 4.

### a)

## 5.


## 6.



## 7.


```{r ex7_import_data}
library(ISLR2)
```

### a)
```{r ex7a}
data <- Boston
head(data)

# Get number of rows
row_number <- nrow(data)

# Get number of columns
col_number <- ncol(data)

```

Number of rows represent the number of observations in our dataset. And number of 
columns represent the number of predictors we have.

### b)
```{r}
plot(data$age, data$crim)
```

